---
title: "Untitled"
author: "Carlos Diaz"
date: "8 de junio de 2018"
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r} 
library(rmarkdown)
library(knitr)

```

# REPRODUCIBLE RESEARCH COURSE - DATA SCIENCE SPECIALIZATION
## ASSIGNMENT: COURSE PROJECT 1  

### 1. Code for reading in the dataset and/or processing the data: 

```{r}
## Reading csv datafile: 
DataActivity <- read.csv("activity.csv")
str(DataActivity)
DataActivityNAs <- na.omit(DataActivity)
str(DataActivityNAs)
DataActivityNAs$date <- as.Date(DataActivityNAs$date, "%Y-%m-%d")
```

### 2. Histogram of the total number of steps taken each day:

```{r}
library(ggplot2)
TotalSteps <- tapply(DataActivityNAs$steps, DataActivityNAs$date, FUN=sum)
hist(TotalSteps, main="Total Steps per Day")
g <- ggplot(DataActivityNAs, aes(date,steps))
g <- g+geom_bar(stat="identity")+labs(title="Total Number of Steps per Day", x= "Date", y="Number of Steps")
print(g)
```

### 3. Mean and median number of steps taken each day:

```{r}
summary(TotalSteps)
MeanSteps <- mean(TotalSteps)
MeanSteps
MedianSteps <- median(TotalSteps)
MedianSteps
```
### 4. Time series plot of the average number of steps taken:
```{r}
AvgDailySteps <- aggregate(steps~interval, data=DataActivityNAs, FUN=mean)
plot1 <- plot(AvgDailySteps, type="l")
```


### 5. The 5-minute interval that, on average, contains the maximum number of steps:

```{r}
IntMaxSteps <- AvgDailySteps$interval[which.max(AvgDailySteps$steps)]
print(IntMaxSteps)  
```

### 6. Code to describe and show a strategy for imputing missing data:

```{r}
## Creating a new dataset that is equal to the original but with the missing data filled in. 
ImputedData <- merge(DataActivity, AvgDailySteps, by= "interval", suffixes = c("", ".y"))
NAvalues <- is.na(ImputedData$steps)
ImputedData$steps[NAvalues] <- ImputedData$steps.y[NAvalues]
ImputedData <- ImputedData[, c(1:3)]
sum(is.na(ImputedData$steps))
ImputedData$date <- as.Date(ImputedData$date)
```

### 7. Histogram of the total number of steps taken each day after missing values are imputed:

```{r}
g2 <- ggplot(ImputedData , aes(date, steps))
plot2 <- g2+geom_bar(stat="identity", color = "black")+labs(title="Imputed Total Number of Steps Taken per Day", x="Date", y="Daily Steps")
print(plot2)
```

### 8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends:

```{r}
ImputedDatadays <- ImputedData
Weekenddays <- weekdays((ImputedDatadays$date)) %in% c("Saturday", "Sunday")
ImputedDatadays$daytype <- "Weekday"
ImputedDatadays$daytype[Weekenddays == TRUE] <- "Weekend"
ImputedDatadays$daytype <- as.factor(ImputedDatadays$daytype)

## Daily average steps by daytype
AvgStepsByDayType <- aggregate(steps ~ interval + daytype, ImputedDatadays, mean)
names(AvgStepsByDayType)[3] <- "Avg_Steps"

library(lattice)

plot3 <- xyplot(Avg_Steps ~ interval | daytype, AvgStepsByDayType, type="l", layout=c(1,2), main = "Weekend vs Weekday Average Number of Steps", xlab="5 min Interval", ylab="Avg Steps taken")
print(plot3)
```
